---
title: "Local adaptation portends climate change winners and losers - DRAFT"
runtitle: "Grayling local adaptation"
documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Heidi E Golden
    affiliation: '1'
    etal: Golden # First author's last name. 
    email: hgolden@goldenecology.com # Indicates corresponding Author
    institution: Golden Ecology LLC
    street: 57 E Weatogue St
    city: Simsbury
    state: CT # only USA, Australia, Canada
    zip: 06070
    country: USA
  - name: Linda A Deegan
    affiliation: '2'
  - name: Mark C Urban
    affiliation: '3'
affiliation:
  - id: '1'
    department: 
    institution: Golden Ecology LLC
    city: Simsbury
    state: CT # only USA, Australia, Canada
    country: USA
  - id: '2' 
    department: 
    institution: Woodwell Climate Research Center
    city: Falmouth
    state: MA # only USA, Australia, Canada
    country: USA
  - id: '3' 
    department: Ecology and Evolutionary Biology Dept.
    institution: University of Connecticut
    city: Storrs
    state: CT # only USA, Australia, Canada
    country: USA
output: rticles::frontiers_article
csl: frontiers.csl
bibliography: test.bib
---

\begin{abstract}
  Arctic freshwater species are highly susceptible to extinction due to rapid polar amplification of climate change and dispersal limitation from polar and dendritic habitat constraints. Plasticity and/or local adaptation (evolution) of traits might help mitigate impacts of climate mediated environmental stress, such as increased water temperature. Arctic grayling (\textit{Thymallus arcticus}) provide a model Arctic freshwater species for investigating trait variation among populations that might mitigate impacts from rapid climate change. We used common garden experiments to compare reaction norms for early life-history traits in response to temperature differences among two Arctic grayling populations that experience different local temperature regimes (cold and warm). We reared Arctic grayling using a sib-ship mating design under three different temperatures (8°C, 12°C, and 16°C). We found evidence for local trait adaptation to temperature suggesting excelerated growth and metabolic rates for the cold-adapted population compared to the warm-adapted populations. However, the cold-adapted population showed reduced survivorship compared to the warm-adapted population under warm, projected climate change, conditions. Our finding suggest that trait evolution by local populations might help mitigate species extinctions due to rapid climate change, but population persistence likely depends on trade-offs between growth and survivorship.

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Climate Arctic Traits Extinction Freshwater Polar Dispersal Grayling } 

\end{abstract}


# Introduction {-}
Arctic freshwater species are highly susceptible to extinction from rapid climate change because they are dispersal-limited by polar range and dendritic habitat constraints [@Rantanen2022;@Reist2006;@Song2021]. However, these dispersal-limited populations might moderate extinction probability if their phenotypic variation and/or genetic adaptive capacity for traits promote survival under new climate change conditions [@Roman-Palacios2020]. Phenotypic trait plasticity (flexibility in trait expression) and local trait adaptation (genotypic evolution) to warmer water temperature might help mitigate the impacts of freshwater species loss due to rapid climate change [@Gunderson2015;@Roman-Palacios2020]. Understanding the limitations and underlying mechanisms of species phenotypic responses to environmental change will enable better estimation and prediction of the magnitude of climate change extinction risk for Arctic freshwater species.

Common garden experiments provide means to assess population-level plasticity and evolutionary capacity by testing for trait differences (reaction norms) among populations reared under the same set of environmental conditions. Evidence from common garden experiments suggests that early life history traits in salmonid fishes, including yolk assimilation and size at hatching, could be under strong selection and appear to be positively correlated with survival [@Miller1988;@Perez2010]. For example, populations of sockeye salmon (\textit{Oncorhnchus nerka}) showed variation in thermal tolerance limits and embryo survival when reared in a common garden setting [@Whitney2013]. Populations of brown trout (\textit{Salmo trutta}) reared in a common garden environment showed evidence of locally adapted early life history traits with implications for climate change adaptability [@Jensen2008]. European grayling (Thymallus thymallus) populations from common garden experiments showed evidence for local adaptation among populations for larval growth, yolk conversion efficiency, and survival [@Haugen2000;@Kavanagh2010;@Thomassen2011], but growth-rate and developmental trade-offs in cold-adapted populations reared at higher temperatures suggested limits to adaptive capacity [@Kavanagh2010]. Thus,the response of Arctic freshwater fish, that are limited in dispersal ability by polar and dendritic habitats, to rapid climate change might hinge on degree of phenotypic plasticity, adaptive potential, and trade-offs among early life history traits.

Arctic grayling (\textit{Thymallus arcticus}), a cold-adapted freshwater salmonid species, on Alaska’s North Slope shows inter-population neutral genetic variation due to population isolation (distance and environment) and downstream-biased dispersal from the headwaters toward the Arctic Ocean [@Golden2021]. North Slope Arctic grayling populations show variation in growth of age-0 fish among streams, which was shown to be related to stream temperature and the presences of multiple lakes within the watershed [@Lueke2008]. However, trait variation either by phenotypic plasticity or natural selection within these semi-isolated populations has not been assessed and might be important for species persistence during rapid climate change. Few common garden experiments have been conducted with Arctic grayling and none have been conducted in the Arctic to test for local trait adaptation. In this study, we investigated the ability of Arctic grayling to respond to future climate change conditions by examining underlying mechanisms for early life history phenotypic trait variation in the Alaskan Arctic, an area undergoing the most rapid rate of climate change on earth.

We investigated two North Slope headwater populations of Arctic grayling that experience different thermal regimes and ask (1) are there differences in early life-history traits (means and reaction norms) among populations and, if so, (2) are differences due to trait plasticity (environment) or local adaptation (genetics), and (3) are trait differences associated with trade-offs with survival?

# Materials and Methods {-}
## Study species and area {-}
Arctic grayling is a freshwater salmonid species with a Holarctic distribution, including northern regions of Europe, North America, and parts of Asia, and with small remnant populations in Montana, USA. The species is threatened by increasing water temperature due to climate change and by other anthropogenic factors, including habitat degradation and fragmentation [@AEP2015;@Tingley2022]. In the Alaskan Arctic, reduced aquatic connectivity due to climate change and the shifting balance between precipitation and evapotranspiration further influences dispersal capability for Arctic grayling [@Betts2015;@Golden2021]. These combined challenges will likely increase extinction probability unless the species is able to adjust to warmer climate change conditions through phenotypic plasticity or adaptive evolution.

```{r Figure1_SiteMap, cgSites, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='The Kupark River and Sagavanirktok River Arctic grayling populaiton collection locations.', out.width = "85mm", out.height = "85mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

Arctic grayling populations used in this research were previously determined to be neutrally genetically differentiated and to exhibit phenotypic differences among populations regarding seasonal movement patterns [@Golden2016;@Golden2021]. Individuals from the Kuparuk River (Kup) population were collected from the Kuparuk River headwaters, where individuals overwinter in the headwater lake (Green Cabin Lake) and migrate downstream in springtime into the Kuparuk River to spawn (Figure 1). Individuals from the Sagavanirktok River (Sag) population were collected from the lower reaches of Oksrukuyik Creek, where individuals overwinter in the Sagavanirktok River and migrate upstream in springtime into Oksrukuyik Creek to spawn (Figure 1). Both populations exhibit moderate to high effective population size (Ne), inbreeding coefficients close to zero (+/- 0.04) and low gene flow among populations (Golden et al. 2021). Additionally, spawning and rearing locations within these two watersheds experience different thermal conditions during the open-water time period from late-May to mid-September (Figure 2). Average temperature during the open-water period (late-May to Mid-September) for the Sag populations was X-Xdegrees C and for the Kuparuk River population was X-Xdegrees C for YYYY to YYYY.

Adult Arctic grayling from these two populations were captured in late May using fyke nets and by angling and held temporarily (XX Days) on site in holding pens until spawning capable. All Arctic grayling captured were assumed to be randomly sampled from the population. Adult grayling were determined to be spawning capable when gentle pressure on the abdomen toward the vent readily expelled gametes. When spawning ready, gametes of males and females were stripped and families were created for each population by crossing each sire captured with two unique dams, yielding 16 families for the Kuparuk River population and 5 families for the Sagavanirktok population. No initial differences in egg size based on yolk volume estimates existed among the populations (yolk volume means +/- SE: Kup = XX +/- XXmm^3; Sag = XX +/- XXmm^3). Our mating design produced F1 individuals of full-siblings, half-siblings, and unrelated individuals for each population, allowing estimation of early life history trait heritability. Eggs from each population and each family were separated into three treatment groups, cool (8°C), warm (12°C), and hot (16°C), incubated to hatching and reared to fry.

## Study design {-}
Three experimental tanks (1 x 3 m) were located at Toolik Field Station (Figure 1). The water for the experiment was sourced from filtered Toolik Lake water, which was similar in chemical composition to that of the populations used in the environment (SUPPLEMENTAL ??). Three temperature regimes were established for the tanks corresponding to mean July and August water temperatures experienced by the Sagavanirktok (low), Kuparuk (warm) and Predicted Climate Change (hot) conditions, respectively. In the laboratory, the eggs from each full-sib family were divided into three lots of ### eggs. Each lot was put into individual chamber within each tanks (Figure 3). All families were represented at all three temperatures, such that all three populations were equally represented in all tanks. Environmental conditions were carefully controlled in all tanks (Supplemental Temperature data). We performed a test for the presence of
tank (Aquarium??) effects for growth rate or survival probability
(arcsine square-root transformed) during the period
of external feeding. Tanks were nested within populations
and analysed for each temperature separately. We
found no tank effect (all P > 0.2). However, the use of
four tanks for each temperature may introduce some
extra, unexplained variance into the model. This variance
is probably absorbed into the sire and dam effects.
Dead individuals (eggs and fry) were registered and
removed daily. At hatching, up to 15 yolk-sac fry were
sampled from each family and conserved in 70% ethanol
for later measurements. All ®sh were conserved for more
than 4 weeks before measuring. Families experiencing
high mortality were not sampled to ensure a suf®cient
number of individuals for the later parts of the experiment.
The same number of fry was sampled at swim-up
(a distinct time period when the fry leave the bottom of
the boxes and shoal in a mid-water position), leaving a
variable number of ®ngerlings in each box (Aursjùen: 5±
86, HaÊrrtjùnn: 13±88, Lesjaskogsvatn: 5±86). In order to
reduce possible density effects during the period of
exogenous growth, the number of individuals in each
box was reduced to less than 50 individuals by randomly
removing excess individuals. We tested for possible
density effects on growth and survival, but no effect
was found (regression analysis: density vs. survival:
r  0.042, N  143, P  0.615; density vs. growth rate:
r  ±0.005, N  143, P  0.952). Carlstein (1995), working
at somewhat higher densities, also did not ®nd any
density effect on grayling growth or survival. Density is,
therefore, not used as a variable in the statistical analyses.
After swim-up, the fry were fed commercial ®sh pellets
(EWOS EST 90, 0-granulate) in excess once a day for a
period corresponding to about 180 degree-days, which
corresponded to 16, 20 and 23 days for the warm,
medium and cold regimes, respectively. Fungal infections
(Saprolegnia sp.) constituted a problem during the experiment,
which increased after the onset of external
feeding. As a result of this, the warm regime experiment
with the HaÊrrtjùnn grayling (all families in the four warm
treatment tanks) had to be terminated after 13 days of
external feeding.
## Cumulative hatching {-}


## Survival {-}


## Fish length and yolk volume {-}


## Respiration {-}


## Heritability {-}



# Results {-}


## Cumulative hatching {-}


## Survival {-}


## Fish length and yolk volume {-}


## Respiration {-}


## Heritability {-}




<!-- Comments should be HTML-style comments -->

## Subsection 1 {-}

You can use `R` chunks directly to plot graphs.  

```{r graph, echo = TRUE, message=FALSE, fig.show = "hide"}
x <- 0:100
set.seed(999)
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
plot(x, y)
```

## Subsection 2 {-}

Frontiers requires figures to be submitted individually, in the same order as
they are referred to in the manuscript. Figures will then be automatically
embedded at the bottom of the submitted manuscript. Kindly ensure that each
table and figure is mentioned in the text and in numerical order. Permission
must be obtained for use of copyrighted material from other sources (including
the web). Please note that it is compulsory to follow figure instructions.
Figures which are not according to the guidelines will cause substantial delay
during the production process.

# Discussion

 (1) are there differences in early life-history traits (means and reaction norms) among populations and, if so, (2) are differences due to trait plasticity (environment) or local adaptation (genetics), and (3) are trait differences associated with trade-offs with survival?

# Disclosure/Conflict-of-Interest Statement {-}

<!--  
Frontiers follows the recommendations by the International Committee of Medical
Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
that all financial, commercial or other relationships that might be perceived by
the academic community as representing a potential conflict of interest must be
disclosed. If no such relationship exists, authors will be asked to declare that
the research was conducted in the absence of any commercial or financial
relationships that could be construed as a potential conflict of interest. When
disclosing the potential conflict of interest, the authors need to address the
following points:

 - Did you or your institution at any time receive payment or services from a
   third party for any aspect of the submitted work?
 - Please declare financial relationships with entities that could be perceived
   to influence, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 - Please declare patents and copyrights, whether pending, issued, licensed
   and/or receiving royalties relevant to the work.
 - Please state other relationships or activities that readers could perceive to
   have influenced, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 -->

The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

# Author Contributions {-}

<!--  

When determining authorship the following criteria should be observed:

 - Substantial contributions to the conception or design of the work; or the
   acquisition, analysis, or interpretation of data for the work; AND
 - Drafting the work or revising it critically for important intellectual
   content; AND
 - Final approval of the version to be published ; AND
 - Agreement to be accountable for all aspects of the work in ensuring that
   questions related to the accuracy or integrity of any part of the work are
   appropriately investigated and resolved.

Contributors who meet fewer than all 4 of the above criteria for authorship
should not be listed as authors, but they should be acknowledged.
(http://www.icmje.org/roles_a.html)

-->

The statement about the authors and contributors can be up to several sentences
long, describing the tasks of individual authors referred to by their initials
and should be included at the end of the manuscript before the References
section.


# Acknowledgments {-}

Funding:

# Supplemental Data 

Supplementary Material should be uploaded separately on submission, if there are
Supplementary Figures, please include the caption in the same file as the
figure. LaTeX Supplementary Material templates can be found in the Frontiers
LaTeX folder

# References

A Frontier article expect the reference list to be included in this section.
To make that happens, the below syntax can be used. This [feature is from Pandoc citeproc](https://pandoc.org/MANUAL.html#placement-of-the-bibliography) which is used with `frontier_article()` to handle the bibliography

::: {#refs}
:::

# Figures {-}

```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

