---
title: "2017 Common Garden Population Site Map"
author: "Heidi E Golden"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

This code creates a site map of the Kuparuk River and Sagavanirktok River adult fish population capture locations for use in the Common Garden journal article.  


```{r InstallMapbox}
#remotes::install_github('qdread/qdrmapbox')
library(qdrmapbox)
```
Go to Mapbox account and create an API access token to access the Raster Tiles API. 
Save the token in a text file, containing only the token and nothing else.
Paste the access token file address below:

```{r AssignAccessToken}
# Read the KML file
set_mapbox_api_key("/Users/heidigolden/Dropbox/00_Consulting/ArcLTER/2023_ArcLTER/Reporting/2023_ArcLTER_Fish_Map/01_data/2023_ArcLTER_Fish_Sites_AccessToken.txt")
```

Download map tiles:
Create bounding box
assign bounding box to an object

```{r DownloadTiles}
zoom <- 12
upper_left <- c(68.880342,	-149.716514)
lower_right <- c(68.492554, -148.647533) 

cg_tile_index <- find_tile_numbers(zoom = zoom, upper_left = upper_left, lower_right = lower_right)

```

Request Mapbox tiles from the API,
Download tiles,
Save tiles to jpg file.

The function download_mapbox_tiles() downloads the tiles into a local directory given the index numbers we just found. Mapbox has low and high resolutions and JPEG quality settings of 70%, 80%, and 90%. I chose the highest in both cases. We saved the output of download_mapbox_tiles() to a data frame called cg_tile_df because it contains the file names of the downloaded tiles which we can use in the next step to georeference.
```{r DownloadSaveTiles}

download_dir <- '/Users/heidigolden/Dropbox/00_Consulting/ArcLTER/2023_ArcLTER/Reporting/2023_ArcLTER_Fish_Map/01_data/cgtiles'

cg_tile_df <- download_mapbox_tiles(tile_numbers_mat = cg_tile_index, download_dir = download_dir, resolution = 'high', jpg_quality = 90)

head(cg_tile_df, 2)
```

Request Mapbox tiles from the API,
Download tiles,
Save tiles to jpg file.

The function download_mapbox_tiles() downloads the tiles into a local directory given the index numbers we just found. Mapbox has low and high resolutions and JPEG quality settings of 70%, 80%, and 90%. I chose the highest in both cases. We saved the output of download_mapbox_tiles() to a data frame called cg_tile_df because it contains the file names of the downloaded tiles which we can use in the next step to georeference.
```{r DownloadSaveTiles}

download_dir <- '/Users/heidigolden/Dropbox/00_Consulting/UConn/CommonGarden/00_CG_Manuscript/04_figures/cg_raster_tiles/'

cg_tile_df <- download_mapbox_tiles(tile_numbers_mat = cg_tile_index, download_dir = download_dir, resolution = 'high', jpg_quality = 90)

head(cg_tile_df, 2)
```


Georeferencing the tiles:

The function georeference_all_tiles() uses GDAL to georeference the tiles by associating the images with geographical locations. GDAL is a powerful system library for working with vector and raster geospatial data. The command gdal_translate turns the JPEGs into GeoTIFF files by telling GDAL the coordinates of the corners of each tile. The command build_virtual_raster() uses GDAL to mosaic the image tiles together into one “virtual raster,” or VRT file. This function internally calls the command gdalbuildvrt and creates a VRT file called cgimage.vrt in the same directory as the .tif files. The VRT file is a small text file that acts like a raster, though it’s only a pointer to the GeoTIFF tiles, so don’t delete the .tif files once the .vrt file is created.

```{r GeoreferenceTiles}
georeference_all_tiles(cg_tile_df)

build_virtual_raster(cg_tile_df, file.path(download_dir, 'cgimage.vrt'))

```

Making the map:
The map is made with ggplot. We use the function annotation_spatial() from the excellent ggspatial package to turn the raster to an “annotation” … so it can be plotted in its true red-blue-green colors but doesn’t affect the fill or color scales of any of the actual data we plot on the map.

Load the packages we will need to make the map:
```{r LoadMapPackages}
library(sf)
library(raster)
library(ggspatial)
library(ggplot2)
library(tidyr)
library(dplyr)
library(stringi)
require(rgdal)
```

Now read in the satellite imagery as a raster stack (it’s a stack because it actually has three layers: the red, green, and blue).

```{r SatImage}
cgimage_raster <- stack(file.path(download_dir, 'cgimage.vrt'))
```

Get data to plot:

Use the sf package to read the kml file. A kml file is the unzipped versions of a kmz file. 
Unzip the zipped package on your computer. 
A kml file can contain multiple layers in a nested format. 
Check to see which layer you want to use and specify it in the st_read function.

## Site Locations
Create a geospatial data frame for use in creating a site map for collection locations for the Kuparuk River and Sagavanirktok Arctic grayling populations used in the 2017 Common Garden experiment.

```{r CreateSpatialDataset}

Site = c("Kup", "Sag")
Latitude = c(68.537169, 68.863069)
Longitude = c(-149.248228, -148.844709)


Sites <- data.frame(Site, Latitude, Longitude)

cg_df <- st_as_sf(x = Sites,                         
           coords = c("Longitude", "Latitude"),
           crs = 4326)
head(cg_df)

```



```{r Plot_cg_Map}

pallet <- scale_fill_manual(values = c("deeppink3", "dodgerblue"))

gg <- ggplot() +
  annotation_spatial(data = cgimage_raster, alpha = 0.9) +
  geom_sf(data = cg_df, aes(fill=Site), colour="white",pch=21, size = 2) +
  pallet +
 coord_sf(crs = 4326) +
  annotate('text', x = Inf, y = -Inf, label = '\u00a9 Mapbox \u00a9 OpenStreetMap', hjust = 1, vjust = -1, color = 'white', size = 2)+
  labs(x="Longitude (WGS84)", y="Latitude",
       title="Arctic Grayling Population Collection Sites (2017)") +
   # spatial-aware automagic scale bar
  annotation_scale(location = "bl",style = "ticks") +
  theme_bw() +
  # spatial-aware automagic north arrow
  annotation_north_arrow(width = unit(.2,"in"), 
                         pad_y = unit(.2, "in"),location = "bl", 
                         which_north = "true")

# Specify the map area beyond the default
gg + coord_sf(xlim = c( -148.844709, -149.248228), ylim = c(68.537169, 68.863069))

#upper_left <- c(68.880342,	-149.716514)
#lower_right <- c(68.492554, -148.647533) 


ggsave("/Users/heidigolden/Dropbox/00_Consulting/UConn/CommonGarden/00_CG_Manuscript/20231219_CommonGarden_Manuscript_Frontiers/20231219_CommonGarden_Manuscript_Frontiers_files/figure-latex/cgSites.pdf")

```

NOTE:  This map is a bit wonky. Needs to be wider and perhaps have rivers identified.

