----
title: "Yolk Volume Estimates"
author: "Heidi Golden"
date: "11/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

CHANGE TEXT FROM LENGTHS TO YOK VOLUMES

Introduction:
We conducted a common garden experiment in 2017, which included estimating yolk volume for eggs collected prior to fertilization and as yolk volume was depleted over time as larval fish developed. We collected a subsample of individuals from each family within each treatment tank, fixed the subsample in 10% formalin, then preserved the subsample in 70% ethanol until the yolksac dimensions could be measured. 

Egg and larval yolk volumes were measured after being preserved in ethanol.



Load the Data:
The data set consists of fish lengths measured over time. We had three treatment groups (T1 = 16 C, T2 = 12 C, and T3 = 16 C) and three populations (locations) per treatment (K or Kup = Kuparuk, O or Oks = Lower Oksrukuyik, and OZ or OksZev = Upper Oksrukuyik). We also have fish that we raised from eggs (reared) and that we caught from the rivers soon after hatching (wild). 

```{r LoadYolkVolumeData}
yvol <- read.csv("/Users/hgolden/Dropbox/00_Consulting/UConn/CommonGarden/z_2023_archive/00_Last_CommonGarden_Folders/02_original_data/Length-YolkVolume/00_CG_LnYolkVolLifeStage_WokingFile_HG_2022.csv")

 yvol$location <- as.factor(yvol$location)
 yvol$days.post.fert <- as.numeric(yvol$days.post.fert)
 yvol$Treatment <- as.factor(yvol$Treatment)
 yvol$family <- as.factor(yvol$family)
 yvol$disposition <- as.factor(yvol$disposition)
 yvol$reared.wild <- as.factor(yvol$reared.wild)
 yvol$measured.by <- as.factor(yvol$measured.by)
 yvol$duplicate.to.remove <- as.factor(yvol$duplicate.to.remove)
 yvol$yolk.vol <- as.factor(yvol$yolk.vol)
 yvol$growth <- as.factor(yvol$growth)
 yvol$ln.at.hatch <- as.factor(yvol$ln.at.hatch)
 yvol$ln.at.swimup <- as.factor(yvol$ln.at.swimup)
 yvol$respiration <- as.factor(yvol$respiration)
 yvol$experiment <- as.factor(yvol$experiment)
 
str(yvol)
```

First, filter out the ethanol preserved eggs and/or fish that had descicated due to sample ethanol evaporation:

The ethanol preservative in a few of the samples evaporated over time, leaving the sample descicated. We measured these individuals anyway and labeled their "disposition" as "dried.ethanol." We will create a standard curve by purposely drying ethanol-preserved fish that we have already measured and re-measuring each individual after descication. We will regress the descicated fish measurements versus ethanol-preserved measurements to create a function that we can use to estimate pre-descication lengths in order to include these fish in our common garden experiment. 

For now, however, we must remove the dried.ethanol fish from the data.

```{r RemoveDriedEthanolFish}
yvol <- subset(yvol, disposition != "dried.ethanol")
str(yvol$disposition)
```


Here I create a dataset containing "reared" fish (not wild) and length data with "live" fish lengths corrected with the regression formula to approximate shrinkage due to ethanol preservation.
This dataset also excludes fish from the OksZev location
```{r subsetData}
data1 <- subset(yvol, reared.wild == "reared" & location != "OksZev" & experiment == "yolksac")
str(data1)

newd <- data1[ c("location", "days.post.fert", "Treatment", "family", "indiv.1to15", "yolk.sac.volume")]
head(newd)


## IF any changes are made to the original data sheet, or within this script
### You MUST adjust this file below in EXCEL prior to continuing with this R Code.

#write.csv(newd, "/Users/hgolden/Dropbox/00_Consulting/UConn/CommonGarden/00_CG_Manuscript/01_derived_data/01_YolkVolumeDataSet-ForR-20221202.csv")
## Open in Excel and duplicate Post-Fert Egg Volumes (days post fert = 0) for all Treatments
# D0 volumnes were prior to assigning treatment groups, so is the same for all treatment.
```

```{r ReloadAdjustedDataChangeCharToFact}

### Reload adjusted data here for graphing:

adj.data <- read.csv("/Users/hgolden/Dropbox/00_Consulting/UConn/CommonGarden/00_CG_Manuscript/01_derived_data/02_Adjusted-YolkVolumeDataSet-ForR-20221202.csv")
library(dplyr)

cols <- c("location", "Treatment", "family", "indiv.1to15")

adj.data <- adj.data %>%
    mutate_at(cols, list(~factor(.)))
adj.data <- as.data.frame(adj.data)
str(adj.data)

```

## Plots of Initial Egg/Yolk Volumes

```{r PlotInitialEggVol}
#install.packages("tidyverse")
#install.packages("hrbrthemes")
#install.packages("viridis")
#install.packages("ggeasy")

# Libraries
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(ggeasy)

#  dataset
eggs <- subset(adj.data, days.post.fert == 0)
str(eggs)

# Plot

eggs %>% 
  ggplot(aes(x=location, y=yolk.sac.volume, fill=location)) + 
  geom_boxplot(width=0.5,lwd=0.5) +
  geom_jitter(width=0.15, size = 0.05) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)) + 
  labs(x = "Population", y = "Volume (mm^3") +
  ggtitle("Initial Egg Volume") +
  ggeasy::easy_center_title()


# Use this to print the plot to the Figures Folder:
# png("/Users/hgolden/Dropbox/00_Consulting/UConn/CommonGarden/00_CG_Manuscript/04_figures/05_CG_InitialYolkVolume_Graph_20221210.png")

# dev.off

```

```{r ANOVA_InitialEggVolume}

# Perform ANOVA for initial yolk volume by location
aov_test <- aov(yolk.sac.volume ~ location, data = eggs)
  summary(aov_test)


```



YOLK VOLUME DATA - Linear model.
```{r lmyvol}
new.lm <- lm(yolk.sac.volume ~ 0 + days.post.fert * location * Treatment, data = adj.data)

summary(new.lm)
```

CREATE SUMMARY STATISTICS:

Function to Summarize the Data:
```{r SummarizeFunction}
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      n = length(x[[col]]), 
      se = sd(x[[col]])/sqrt(length(x[[col]])), na.rm=TRUE)
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}
```



Summarize the Data:

```{r SummarizeData}

sum.data <- data_summary(adj.data, varname = "yolk.sac.volume", 
                    groupnames = c("location", "days.post.fert", "Treatment"))

#sum.data$days.post.fert <- as.numeric(sum.data$days.post.fert)
sum.data
#write.csv(sum.data, "/Users/hgolden/Dropbox/00_Consulting/UConn/CommonGarden/00_CG_Manuscript/01_derived_data03_CG_SumStats_yvol_20221201.csv")
```


Plot Length Over Time (with "live" converted to "ethanol" estimated length):

```{r Plot_reared}
library(ggplot2)

p <- ggplot(data = sum.data, aes(x = days.post.fert, y = yolk.sac.volume, color = location)) + xlim(0, 40) + ylim(0,25.5) + geom_point() + geom_line() + geom_errorbar(aes(ymin=yolk.sac.volume-se, ymax=yolk.sac.volume+se), width=.2, )

p +  facet_wrap(~ Treatment, labeller = labeller(Treatment = function(string) paste(string, "°C"))) + labs(title = "Yolk Volume", y = "Yolk Volume (mm^3)", x = "Days Post Fertilization") +
  theme_minimal()

```




Analyze the Data:
```{r ANOVA_EstLive}
aov.est <- aov(yolk.sac.volume ~ days.post.fert + location * Treatment, data = newd)
summary(aov.est)
```

