---
title: "Life Stage Filter for Manuscript"
author: "Heidi E Golden"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This file filters the raw data entry file "00_CG_LifeStage_Master_2019_WorkingFile.csv" to include only the Kup (Kuparuk River) and Oks3 (Sagavanirktok River) Populations and only up to DAY 40 Post Fertilization. Kup and Oks3 were the only viable eggs/populations used in the experiment. Data collected post Day 40 did not include enough families to provide valid comparisons of the data.

```{r cars}
data <- read.csv("/Users/heidigolden/Dropbox/00_Consulting/UConn/CommonGarden/00_CG_Manuscript/00_raw_data/LifeStage/00_CG_LifeStage_Master_2019_WorkingFile_HG2022.csv")
str(data)
```

## Subset data to include only Kup and Oks3 Populations

Here we subset the data frame to include the two populations used in the common garden experiment.

This code checks the data frame for unique values for Location, which here is equivalent to "population." 

```{r SubsetToKupOks3}
#Check to make sure all "Location" names were entered correctly
unique(data$Location)

```

Here, we subset the data to include our two common garden populations: Kup and Oks3; i.e., no "wild" fish "WOZ"   "WK"    "WO"

```{r subsetToKupOks3}
# Subsets the data to not include "WOZ"   "WK"    "WO"

#first change the data structure from character to factor
data$Location <- as.factor(data$Location)

# next subset the data frame by "Location"
sub_data <- subset(data, Location %in% c("Kup","Oks 3"))

# View the subsetted dataframe
#tail(sub_data)



```

Next we subset the dataframe to exclude days-post-fertilization beyond Day 40.
```{r SubsetOnlyUpToDay40}
ls_data <- sub_data[sub_data$days.post.fert <= 40, ]

```

Finally, we write this data to a new folder specifically for "clean" data.

```{r WriteFilteredData}
write.csv(ls_data, "/Users/heidigolden/Dropbox/00_Consulting/UConn/CommonGarden/00_CG_Manuscript/02_clean_data/CG_LifeStage_data_clean_2023.csv")
```